---
title: "Schema of Study Designs"
format: 
  revealjs:
      slide-number: true
      logo: images/logo.png
author: "Dr Abhijit P Pakhare"
institute: "All India Institute of Medical Sciences, Bhopal"
license: CC BY-NC
editor: visual
---

## Identify the design

```{r}
library(DiagrammeR)
grViz("
digraph generic {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]
H [label = '@@8', fillcolor= LightPink]
I [label = '@@9', fillcolor= LightPink]
# several 'edge' statements
A->B A->C B->D B->E C->F
C->G 
edge [arrowhead = none, arrowtail = none, style=tapered, label= 'estimate', fontsize=6]
D->H E->H F->I G->I
}
  [1]: ''
  [2]: ''
  [3]: ''
  [4]: ''
  [5]: ''
  [6]: ''
  [7]: ''
  [8]: ''
  [9]: ''
  ")
```

## Design of randomized controlled trial (1/3)

```{r}
library(DiagrammeR)
grViz("
digraph rct1 {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]

# several 'edge' statements
edge [label= 'randomization', fontsize=6]

A->B A->C 


}
  [1]: 'Population'
  [2]: 'Intervention'
  [3]: 'Placebo or\\nStandard Care'
  ")
```

## Design of randomized controlled trial (2/3)

```{r}
library(DiagrammeR)
grViz("
digraph rct2 {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]

# several 'edge' statements
edge [label= 'randomization', fontsize=6]

A->B A->C 

edge [label= 'follow-up', fontsize=6]

B->D B->E 
C->F C->G 

}
  [1]: 'Population'
  [2]: 'Intervention'
  [3]: 'Placebo or\\nStandard Care'
  [4]: 'Meets end point \\nor Outcome'
  [5]: 'Do not meet end point \\nor Outcome'
  [6]: 'Meets end point \\nor Outcome'
  [7]: 'Do not meet end point \\nor Outcome'

  ")
```

## Design of randomized controlled trial (3/3)

```{r}
library(DiagrammeR)
grViz("
digraph rct {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]
H [label = '@@8', fillcolor= LightPink]
I [label = '@@9', fillcolor= LightPink]
# several 'edge' statements
edge [label= 'randomization', fontsize=6]

A->B A->C 

edge [label= 'follow-up', fontsize=6]

B->D B->E 
C->F C->G 
edge [arrowhead = none, arrowtail = none, style=tapered, label= 'estimate', fontsize=6]
D->H E->H F->I G->I
}
  [1]: 'Population'
  [2]: 'Intervention'
  [3]: 'Placebo or\\nStandard Care'
  [4]: 'Meets end point \\nor Outcome'
  [5]: 'Do not meet end point \\nor Outcome'
  [6]: 'Meets end point \\nor Outcome'
  [7]: 'Do not meet end point \\nor Outcome'
  [8]: 'Rate of event or outcome \\nin Intervention'
  [9]: 'Rate of event or outcome \\nin Placebo or SOC'
  ")
```

## Design of cohort study (1/3)

```{r}
library(DiagrammeR)
grViz("
digraph cohort1 {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]

# several 'edge' statements
edge [label= 'ascertain', fontsize=6]

A->B A->C 


}
  [1]: 'Population'
  [2]: 'Exposed'
  [3]: 'Non-exposed'
  
  ")
```

## Design of cohort study (2/3)

```{r}
library(DiagrammeR)
grViz("
digraph cohort2 {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]

# several 'edge' statements
edge [label= 'ascertain', fontsize=6]

A->B A->C 

edge [label= 'follow-up', fontsize=6]

B->D B->E 
C->F C->G 

}
  [1]: 'Population'
  [2]: 'Exposed'
  [3]: 'Non-exposed'
  [4]: 'Incident Cases'
  [5]: 'Do not develop disease'
  [6]: 'Incident Cases'
  [7]: 'Do not develop disease'
  
  ")
```

## Design of cohort study (3/3)

```{r}
library(DiagrammeR)
grViz("
digraph cohort {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]
H [label = '@@8', fillcolor= LightPink]
I [label = '@@9', fillcolor= LightPink]
# several 'edge' statements
edge [label= 'ascertain', fontsize=6]

A->B A->C 

edge [label= 'follow-up', fontsize=6]

B->D B->E 
C->F C->G 
edge [arrowhead = none, arrowtail = none, style=tapered, label= 'estimate', fontsize=6]
D->H E->H F->I G->I
}
  [1]: 'Population'
  [2]: 'Exposed'
  [3]: 'Non-exposed'
  [4]: 'Incident Cases'
  [5]: 'Do not develop disease'
  [6]: 'Incident Cases'
  [7]: 'Do not develop disease'
  [8]: 'Incidence of disease \\nin Exposed'
  [9]: 'Incidence of disease \\nin Non-exposed'
  ")
```

## Design of case control study (1/4)

```{r}
library(DiagrammeR)
grViz("
digraph case_control1 {
graph[rankdir = RL]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]

edge [label= 'select', fontsize=6]

# several 'edge' statements
A->B A->C 
}
  [1]: 'Population'
  [2]: 'Cases'
  [3]: 'Non-Cases'
  
  ")
```

## Design of case control study (2/4)

```{r}
library(DiagrammeR)
grViz("
digraph case_control2 {
graph[rankdir = RL]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]

# several 'edge' statements

edge [label= 'select', fontsize=6]

A->B A->C 

edge [label= 'assess', fontsize=6]

B->D B->E 
C->F C->G 
}
  [1]: 'Population'
  [2]: 'Cases'
  [3]: 'Non-Cases'
  [4]: 'Exposed'
  [5]: 'Non-Exposed'
  [6]: 'Exposed'
  [7]: 'Non-Exposed'
 
  ")
```

## Design of case control study (3/4)

```{r}
library(DiagrammeR)
grViz("
digraph case_control3 {
graph[rankdir = RL]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]
H [label = '@@8', fillcolor= LightPink]
I [label = '@@9', fillcolor= LightPink]
# several 'edge' statements
edge [label= 'select', fontsize=6]

A->B A->C 

edge [label= 'assess', fontsize=6]

B->D B->E 
C->F C->G 

edge [arrowhead = none, arrowtail = none, style=tapered, label= 'estimate', fontsize=6]
D->H E->H 
F->I G->I
}
  [1]: 'Population'
  [2]: 'Cases'
  [3]: 'Non-Cases'
  [4]: 'Exposed'
  [5]: 'Non-Exposed'
  [6]: 'Exposed'
  [7]: 'Non-Exposed'
  [8]: 'Odds of Exposure \\nin Cases'
  [9]: 'Odds of Exposure \\nin Non-Cases'
  ")
```

## Design of case control study (4/4)

```{r}
library(DiagrammeR)
grViz("
digraph case_control {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]
H [label = '@@8', fillcolor= LightPink]
I [label = '@@9', fillcolor= LightPink]
# several 'edge' statements
edge [label= 'select', fontsize=6]

A->B A->C 

edge [label= 'assess', fontsize=6]

B->D B->E 
C->F C->G 

edge [arrowhead = none, arrowtail = none, style=tapered, label= 'estimate', fontsize=6]
D->H E->H 
F->I G->I
}
  [1]: 'Population'
  [2]: 'Cases'
  [3]: 'Non-Cases'
  [4]: 'Exposed'
  [5]: 'Non-Exposed'
  [6]: 'Exposed'
  [7]: 'Non-Exposed'
  [8]: 'Odds of Exposure \\nin Cases'
  [9]: 'Odds of Exposure \\nin Non-Cases'
  ")
```

## Design of cross-sectional( analytical type)

```{r}
library(DiagrammeR)
grViz("
digraph cross_sectional {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]
H [label = '@@8', fillcolor= LightPink]
I [label = '@@9', fillcolor= LightPink]
# several 'edge' statements
edge [label= 'classify', fontsize=6]

A->B A->C 

edge [label= 'assess', fontsize=6]

B->D B->E 
C->F C->G 

edge [arrowhead = none, arrowtail = none, style=tapered, label= 'estimate', fontsize=6]
D->H E->H 
F->I G->I
}
  [1]: 'Population'
  [2]: 'Exposed'
  [3]: 'Non-exposed'
  [4]: 'Presence of outcome'
  [5]: 'Absence of outcome'
  [6]: 'Presence of outcome'
  [7]: 'Absence of outcome'
  [8]: 'Prevalence of outcome \\nin Exposed'
  [9]: 'Prevalence of outcome \\nin Non-exposed'
  ")
```

## Identify the design

```{r}
library(DiagrammeR)
grViz("
digraph generic {
graph[rankdir = LR]

node [fontname = Helvetica, shape=box,  style=filled]

A [label = '@@1', fillcolor= LightBlue]
B [label = '@@2', fillcolor= SandyBrown]
C [label = '@@3', fillcolor= LimeGreen]
D [label = '@@4', fillcolor= Salmon]
E [label = '@@5', fillcolor= LightCyan]
F [label = '@@6', fillcolor= Salmon]
G [label = '@@7', fillcolor= LightCyan]
H [label = '@@8', fillcolor= LightPink]
I [label = '@@9', fillcolor= LightPink]
# several 'edge' statements
A->B A->C B->D B->E C->F
C->G 
edge [arrowhead = none, arrowtail = none, style=tapered, label= 'estimate', fontsize=6]
D->H E->H F->I G->I
}
  [1]: ''
  [2]: ''
  [3]: ''
  [4]: ''
  [5]: ''
  [6]: ''
  [7]: ''
  [8]: ''
  [9]: ''
  ")
```

# Thank You
